Measurement Bias from Address Aliasing

### 2016-05-15 presentation.tex
wip: Write first half of presentation

\documentclass{beamer}

\title{Measurement Bias from Address Aliasing}
\subtitle{How your user name affects runtime performance}

@@

\begin{document}
  \frame{\titlepage}

  \begin{frame}
    \frametitle{Measurement Bias}

@@

@@

    % Image from Master's (redo in Tikz probably) showing exact address
    % collisions. Update with addresses taken from paper.

    % Maybe show disassembly with read/write false conflicts.

@@

    % Some points on hardware perf counters.

@@

>>>>

\documentclass{beamer}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{mathtools}
\usepackage{tikz}

\usetikzlibrary{patterns,shapes,positioning,calc}
\setbeamertemplate{bibliography item}[text]
\lstset{language=C, basicstyle=\ttfamily\scriptsize, columns=flexible, belowskip=0pt, aboveskip=-1pt}

\title{Measurement Bias from Address Aliasing}

@@

\begin{document}

\frame{\titlepage}

\begin{frame}

\frametitle{Outline}

\begin{itemize}
  \item What is measurement bias?
  \item What is address aliasing?
  \item How does this help?
\end{itemize}

\begin{frame}[fragile, allowframebreaks]
\frametitle{Measurement Bias}

@@

\begin{itemize}
  \item Runtime performance change from external factors
  \begin{itemize}
    \item Environment variables
    \item Link ordering
  \end{itemize}
  \item Challenge for performance analysis
\end{itemize}

\begin{figure}
  \begin{lstlisting}[frame=single, xleftmargin=.01\textwidth, xrightmargin=.01\textwidth]
static int i, j, k;

int main() {
    int g = 0, inc = 1;
    for (; g < 65536; g++) {
        i += inc;
        j += inc;
        k += inc;
    }
    return 0;
}
  \end{lstlisting}
  \caption{\label{fig:microkernel}Microkernel first presented by Mytkowicz \emph{et al.}\cite{Mytkowicz:2009:WrongData}, showing bias to environment size.}
\end{figure}

What is the execution time of this function?

\pgfplotstableread{bin/microkernel-cycles-haswell.dat}{\stackoffsettable}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
        width = \textwidth,
        height = 5.0cm,
        font = \footnotesize,
        xlabel=Bytes added to environment,
        ylabel=Cycles,
        domain = 0:8192,
        xtick = {0,1024,...,8192},
        xmin = 0,
        xmax = 8192,
        cycle list name = exotic
      ]
      \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \stackoffsettable ;
    \end{axis}
  \end{tikzpicture}
  \caption{Measured average of 10 cycle count samples for 512 different environments; offseting stack by multiples of 16 bytes.}
\end{figure}

@@

\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}

\begin{itemize}
  %\item Collisions between static and automatic variables, depending on alignment of stack.
  \item Address of i, j and k are fixed for all environments.
  \item Address of g and inc is pushed down with length of environment variables.
  \item Spike when inc is at 0x7fffffffe03c + $ k \times $ 0x1000, \emph{aliasing} with i in the last 12 bits.
  \item Performance impact from \emph{false dependencies} in CPU out of order execution.
  \item Effect known as \emph{address aliasing}.
\end{itemize}

\end{column}
\hfill
\begin{column}{.48\textwidth}

% Virtual memory/randomization illustration
\begin{figure}
  \begin{tikzpicture}[font=\footnotesize]
    % See page 453 in pgf manual
    \node [
      rectangle split, rectangle split parts=5, 
      rectangle split part fill={white, white, lightgray, white, lightgray},
      draw, anchor=center, text width=2.5cm
    ] (m) {
        \nodepart{one}
          \scriptsize{USER=lars;SHELL
          =/bin/bash;PWD=
          /home (...)}
        \nodepart{two}
          .stack
          \begin{flushright}
          inc \\
          g
          \end{flushright}
        \nodepart{four}
          .bss
          \begin{flushright}
          k \\
          j \\
          i
          \end{flushright}
      };

      \node[right, yshift=0cm] at (m.two east) {
        \scriptsize{0x7fff'ffffe{\color{red}03c}} % inc
      };
      \node[right, yshift=-0.4cm] at (m.two east) {
        \scriptsize{0x7fff'ffffe038} % g
      };

      \node[right, yshift=0cm] at (m.four east) {
        \scriptsize{0x601044} % k
      };
      \node[right, yshift=-0.4cm] at (m.four east) {
        \scriptsize{0x601040} % j
      };
      \node[right, yshift=-0.8cm] at (m.four east) {
        \scriptsize{0x601{\color{red}03c}} % i
      };

    %\draw [decorate, decoration={brace, amplitude=5pt}] (m.south west) -- (m.seven split west) 
    %  node [black, midway, xshift=-1.5cm, text width=2.0cm] 
    %    { Program code and static data } ;

    %\node[right] at (m.north east) { \scriptsize{0x7fff'ffffffff} } ;
    %\node[right] at (m.south east) { \scriptsize{0x400000} } ;
  \end{tikzpicture}
  \caption{4K address aliasing between stack and static data.}
\end{figure}

% Note: Could have gotten more collisions if g also was colliding with something,
%       but not noticeable effect on cycles executed.

\end{column}
\end{columns}

@@

  % Some points on hardware perf counters.
  \begin{itemize}
    \item Using processor-specific instrumentation, we can \emph{count} various low-level events during execution.
    \item Intel manual lists hundreds of events.
  \end{itemize}

  \begin{description}
    \item[{\small LD\_BLOCKS\_PARTIAL.ADDRESS\_ALIAS.}] 
    \emph{``Counts the number of loads that have partial address match with preceding stores, causing the load to be reissued.''} 
    \cite[B.3.4.4]{OptimizationManual}
  \end{description}
  %This event is listed in the manual for microarchitectures going back to ``Nehalem'', including ``Ivy Bridge'' and ``Haswell'' CPUs~\cite{Volume3B}.
  %Older architectures such as Core do not have this particular event listed, but 4K aliasing is covered by a more general event:
  \begin{description}
    \item[{\small LOAD\_BLOCKS.OVERLAP\_STORE.}] 
    \emph{``Loads that partially overlap an earlier store, or 4-Kbyte aliased with a previous store.''} 
    \cite[Table 19-17]{Volume3B}
  \end{description}

@@

\begin{frame}

\frametitle{References}

\scriptsize{\bibliographystyle{IEEEtran}}
\bibliography{references}
